---
title: 'Lesson 9: Chapter 27 R Markdown and Git in R'
output:
  pdf_document: default
  html_document:
    df_print: paged
  html_notebook: default
date: 2022-03-29
---

# 0. Load libraries

```{r setup, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  comment = "#>",
  collapse = TRUE
)

library(tidyverse)
library(rlang)

# A library for string operations
library(stringr)

# A library for working with Categorical variables, i.e. factors
library(forcats)

# A library for working with date and time
library(lubridate)

# Load NYC flight dataset 
library(nycflights13)
```

```{r}
x <- 1
```

# 1. R Markdown

Markdown is a framework for combining code, results, comments, documentation, etc. in one place. It is fully reproducible. It can be used to write slides, web pages, blogs, and even books. The output can be PDF, MS Word, html.

Here is a good book to learn R MarkDown: [**R Markdown: The Definitive Guide**](https://bookdown.org/yihui/rmarkdown/).


## Start a R Markdown file

Structure of a R Markdown file

* **YAML header**: tells R how to compile the whole file

* **Chunks**: where your R code goes.

* **Normal Text**: where your thinking goes. Rich format with almost everything from Greek letters to calculus formulas. 

$x_2 = y^2$


## Text format

* Bold, italic, 

* Headings

* Lists

## Chunks options

* echo

* message

* include

# 2. Using Git in R Studio

* What is Git?

  * How I collaborate with others in coding before
  
  * How I do it now using Git
  
A good book: [*Happy Git and GitHub for the useR*](https://happygitwithr.com/)






















