---
title: "Exercise solution of lesson 5"
author: "Hao Wang"
date: "3/8/2022"
output:
  pdf_document: default
  html_document: default
---

# 0. Load libraries

```{r setup, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(rlang)

library(readxl)

```

# 4. Exercise

## 4.1 Exercise 1.

- Read the sheet **PBI New** and **SAS New** of Excel file: **Test PBI and SAS new and endorsement 20220214.xlsx** into R.

Solution: 
```{r }
PBI_new <- 
  readxl::read_excel(
    path = "L:/BIRA/R Training/Test PBI and SAS new and endorsement 20220214.xlsx",
    sheet = "PBI New"
    ) %>% 
  as_tibble()

str(PBI_new)

SAS_new <- readxl::read_excel(
    path = "L:/BIRA/R Training/Test PBI and SAS new and endorsement 20220214.xlsx",
    sheet = "SAS New"
    ) %>% 
  as_tibble()

str(SAS_new)
```

- How many records in each tab? 

```{r}
nrow(PBI_new)
nrow(SAS_new)
```

- what's the min, max, and mean of column "DateDiff" in sheet **PBI New** and column "Bus_Days_End_to_End" in sheet **SAS New**.

```{r}
PBI_new %>% summarise(min_DateDiff = min(DateDiff),
                   max_DateDiff = max(DateDiff),
                   mean_DateDiff = mean(DateDiff),
                   )

SAS_new %>% summarise(min_Bus_Days_End_to_End = min(Bus_Days_End_to_End, na.rm = T),
                   max_Bus_Days_End_to_End = max(Bus_Days_End_to_End, na.rm = T),
                   mean_Bus_Days_End_to_End = mean(Bus_Days_End_to_End, na.rm = T),
                   )

```


## 4.2 Exercise 2.

- Load the Denodo view bi_crm.Transfile_Company_Table_2022 in R.

```{r}
library(odbc)

# setup the connection
con <- dbConnect(odbc::odbc(), "DenodoODBC", timeout = 30)

# query table using SQL
transfile_company_table <- 
  dbGetQuery(con, 'SELECT * FROM bi_crm."Transfile_Company_Table_2022"') %>% 
  as_tibble()
```

- sum of "Small Business Financial Customer Served Partnership Count", and sum of "Medium Financial Customer Served Count"

```{r}
transfile_company_table %>% 
  summarise(
    `Total Small Business Financial Customer Served Partnership Count` = 
      sum(`Small Business Financial Customer Served Partnership Count`, na.rm = T),
    `Total Medium Financial Customer Served Count` = 
      sum(`Medium Financial Customer Served Count`, na.rm = T))

  
```



```{r}

```








