---
title: 'Lesson 5: Chapter 10 and 11'
author: "Hao Wang"
date: "3/1/2022"
output:
  pdf_document: default
  html_document: default
---

# 0. Load libraries

```{r setup, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(rlang)

# Load NYC flight dataset 
library(nycflights13)
```

# 1. Data wrangling

Import -> Tidy -> Transform


# 2. Chapter 10. Tibbles: A new way of data frame in R

How to know if an object is a data frame or a tibble?

```{r}
str(iris)
```

```{r}
class(iris)
```


```{r}
str(as_tibble(iris))
```

```{r}
class(as_tibble(iris))
```

- tibbles are more strict.


# 3. Chapter 11. Data import

How to load excel file or csv file into R? 

- **readr** package in tidyverse.

  function: read_csv()
  
- **readxl** package for importing .xls and xlsx

```{r}
library(readxl)

weekly_plan <- 
  readxl::read_xlsx(path = "weekly time plan.xlsx") %>% 
  select(-`Preparation need from instructor`)
```



How to connect to Denodo through ODBC? 

After DenodoODBC is installed, you can use library **odbc** to query Denodo views.

```{r, eval=FALSE}
library(odbc)

# setup the connection
con <- dbConnect(odbc::odbc(), "DenodoODBC", timeout = 30)

# query table using SQL
transfile <- 
  dbGetQuery(con, 'SELECT * FROM bi_corporate_results."Transfile_2022_Final"') %>% 
  as_tibble()

```


# 4. Exercise

## 4.1 Exercise 1.

- Read the sheet **PBI New** and **SAS New** of Excel file: **Test PBI and SAS new and endorsement 20220214.xlsx** into R.

- How many records in each tab? 

- what's the min, max, and mean of column "DateDiff" in sheet **PBI New** and column "Bus_Days" in sheet **SAS New**.

## 4.2 Exercise 2.

- Load the Denodo view bi_crm.Transfile_Company_Table_2022 in R.

- sum of "Small Business Financial Customer Served Partnership Count", and sum of "Medium Financial Customer Served Count"











