---
title: 'Lesson 8: Chapter 16 Dates and Time'
output:
  html_document:
    df_print: paged
  pdf_document: default
  html_notebook: default
---

# 0. Load libraries

```{r setup, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(rlang)

# A library for string operations
library(stringr)

# A library for working with Categorical variables, i.e. factors
library(forcats)

# A library for working with date and time
library(lubridate)

# Load NYC flight dataset 
library(nycflights13)
```

# 1. Working with Date and Time in Tidyverse

**lubridate** is the package in tidyverse to handle all Date and Time related work. Here is the link to package [*lubridate*](https://lubridate.tidyverse.org/) 

### Useful functions:

- today(): `r today()`

- now(): `r now()`

### Creating date, date-time from a string:

```{r}
ymd("2017-01-31")
dmy("31-Jan-2017")
mdy("Jan 31, 2017")
```

```{r}
date_1 <- ymd("2017-01-31")
class(date_1)
```

Date objects are stored in R as integer values, allowing for dates to be compared and manipulated as you would a numeric vector. Logical comparisons are a simple.

```{r}
date_num <- as.numeric(date_1)
date_num
as_date(date_num #, origin = "1970-01-01"
        )
```

### Getting components from a date-time

- functions: year(), month(), mday(), yday(), wday(), hour(), minute(), second()

### Rounding date

- functions: floor_date(), ceiling_date()

```{r}
# create a funcion
make_datetime_100 <- function(year, month, day, time) {
  make_datetime(year, month, day, time %/% 100, time %% 100)
}

flights_dt <- flights %>% 
  filter(!is.na(dep_time), !is.na(arr_time)) %>% 
  mutate(
    dep_time = make_datetime_100(year, month, day, dep_time),
    arr_time = make_datetime_100(year, month, day, arr_time),
    sched_dep_time = make_datetime_100(year, month, day, sched_dep_time),
    sched_arr_time = make_datetime_100(year, month, day, sched_arr_time)
  ) %>% 
  select(origin, dest, ends_with("delay"), ends_with("time"))

head(flights_dt)
```

The function floor_date return a date of the starting day of the week, the default starting day is Sunday.

```{r}
flights_dt %>% mutate(week = floor_date(dep_time, "week")) %>% head()
```
Use the count and floor_date to count the number of flight by week.
```{r}
flights_dt %>% 
  count(week = floor_date(dep_time, "week")) %>% 
  ggplot(aes(week, n)) +
    geom_line()
```

### Time Zone

```{r}
Sys.timezone()
```


# 2. A Quick Summary about what we have learned

- Chapter 2 data visualization: ggplot2

- Chapter 5 data transformation: dplyr

- Chapter 11 data import: readr, odbc, readxl

- Chapter 12 data transformation: rlang

- Chapter 13 data join: 

- Chapter 14 string functions: stringr

- Chapter 15 Categorical variable: forcats

- Chapter 16 Date and Time: lubridate

