---
title: "Lesson_3: Chapter 5 Data transformation"
author: "Hao Wang"
date: "1/26/2022"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(rlang)

# Load NYC flight dataset 
library(nycflights13)
```

# 1. Review Chapter 3. 

#### $3.3.1 Exercises

1. 

```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, color = "blue"))
```

Ans: The error is "blue" is not a column in mpg. You need to put color outside of aes.

```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy), color = "blue")
```

3. Map a continuous variable to color, size, and shape. How do these aesthetics behave differently for categorical vs. continuous variables?

Ans: Continuous variables can be mapped to _color_ and _size_, only categorical variables can be mapped to _shape_.


6. What happens if you map an aesthetic to something other than a variable name, like aes(colour = displ < 5)? Note, you’ll also need to specify x and y.

```{r}
ggplot(mpg, aes(x = displ, y = hwy, colour = displ < 5)) +
  geom_point()
```

#### $3.5.1 Exercises

1. What happens if you facet on a continuous variable?

Ans: The continuous variable is converted to a categorical variable, and the plot contains a facet for each distinct value.

# 2. Chapter 5 Data transformation

We will learn how to use the package __dplyr__. 

-Key functions in **dplyr**:

1. filter(): selct rows.

2. select(): select columns.

3. arrange(): sort by.

4. mutate(): create new columns.

5. summmarise(): usually used with group_by() to create summarized tables by columns.


Explore data set flights in nycflights13: 
```{r}
?flights
str(flights)
```

## 2.1 filter()

```{r}
filter(flights, day == 1, dep_delay >= 180)
```

## 2.2 comparison operators and logical operators, NA

==, !=, etc.

&, |, ! 

%in%

is.na()

## 2.3 arrange()

```{r}

flights_filtered <- filter(flights, day == 1, dep_delay >= 180)
arrange(flights_filtered, month, dep_delay)

```

## 2.4 select()

```{r}
select(flights, year, month, day)
```

There are number of ways to select columns. It can be very useful when you have many columns.

## 2.5 mutate()

```{r}
mutate(flights_filtered, 
       gain = dep_delay - arr_delay,
       speed = distance / air_time * 60
       )
```

A lot of functions can be used in mutate() to create new columns.

## 2.6 summarise() with group_by()

Mean of dep_delay.
```{r}
summarise(flights, delay = mean(dep_delay, na.rm = TRUE))
```
*Question*: what would happen if na.rm = TRUE is removed?

## 2.7 Use the pipe to connect operations

The pipe **%>%** is a very useful operator in tidyverse.  

```{r}
delays <- flights %>% 
  group_by(dest) %>% 
  summarise(
    count = n(),
    dist = mean(distance, na.rm = TRUE),
    delay = mean(arr_delay, na.rm = TRUE)
  ) %>% 
  filter(count > 20, dest != "HNL")

delays
```

Explore the summary functions in R.

You can group by multiple variables. 

Here is the summary table for number of flights every month, day.
```{r,message=FALSE}
daily <- flights %>% 
  group_by(year, month, day) %>% 
  summarise(flights = n())
```

```{r}
daily
```

group_by() can also be used with mutate() and filter().

# 3. Exercises

$5.2.4: 1, 3

$5.5.2: 4

$5.6.7: 4. 5.

$5.7.1: 2. 3. 6. 8























































