---
title: 'Lesson 7: Chapter 14 & 15'
output:
  html_notebook: default
  pdf_document: default
---

# 0. Load libraries

```{r setup, message=TRUE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(rlang)

# A library for string operations
library(stringr)

# A library for working with Categorical variables, i.e. factors
library(forcats)


# Load NYC flight dataset 
library(nycflights13)
```

Here is the link to package [*stringr*](https://stringr.tidyverse.org/) and package [*forcats*]( https://forcats.tidyverse.org/). 

# 1. Chapter 14 Strings

*stringr* has a lot of functions on string manipulation. Function usually starts with str_

- combining strings: str_c()

- spliting strings: str_split()

- subsetting strings: str_sub()

- detect matches: str_detect()

- replacing matches: str_replace()


```{r}
str_c("x", "y", sep = ", ")
```

```{r}
x <- c("Apple", "Banana", "Pear")
str_sub(x, 1, 3)
```

```{r}
x <- c("apple", "banana", "pear")
str_detect(x, "e")
```

- matching strings with **regular expressions**

Section \$14.3 has a brief introduction about **regular expressions**, which is widely used in many other languages, such as SQL, Python, etc. If you are not familiar with it, I'd recommend read section \$14.3 and practice it.


# 2. Chapter 15 Factors

In most of cases, we'd like to convert character variables/categorical variables into *Factor* in R.

```{r}
x1 <- c("Dec", "Apr", "Jan", "Mar")

sort(x1)
```

We can set up a level vector.

```{r}
month_levels <- c(
  "Jan", "Feb", "Mar", "Apr", "May", "Jun", 
  "Jul", "Aug", "Sep", "Oct", "Nov", "Dec"
)
```

Then create a factor from x1 and the levels.
```{r}
y1 <- factor(x1, levels = month_levels)
y1

sort(y1)

```

To get familar with gss_cat survey data.
```{r}
str(gss_cat)
?gss_cat
```

Use levels() or count() to see the levels in a factor.
```{r}
levels(gss_cat$race)
levels(gss_cat$rincome)

gss_cat %>% count(rincome)
```

## 2.1 Modifying factor order in ggplot2

Explore the average number of hours spent watching TV per day across religions:
```{r}
relig_summary <- gss_cat %>%
  group_by(relig) %>%
  summarise(
    age_mean = mean(age, na.rm = TRUE),
    tvhours_mean = mean(tvhours, na.rm = TRUE),
    n = n()
  )

View(relig_summary)

ggplot(relig_summary, aes(tvhours_mean, relig)) + 
  geom_point()
```

```{r}
ggplot(relig_summary, aes(tvhours_mean, fct_reorder(relig, tvhours_mean))) +
  geom_point() +
  labs(x = "Mean TV hours", y = "Religon")
```

Number of person in each marital status in bar chart. 

```{r}
gss_cat %>%
  ggplot(aes(marital)) +
  geom_bar()
```

Make the bars in an upward order.
```{r}
gss_cat %>%
  mutate(marital = marital %>% 
           fct_infreq() %>% 
           fct_rev()
         ) %>%
  ggplot(aes(marital)) +
    geom_bar()
```































