---
title: "Lesson 6 Chapter 12 & 13"
author: "Hao Wang"
date: "3/9/2022"
output:
  pdf_document: default
  html_document: default
---

# 0. Load libraries

```{r setup, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(rlang)

# Load NYC flight dataset 
library(nycflights13)
```

# $12.3 Pivoting

- pivot_longer(data, cols, names_to = "name", values_to = "value")

```{r}
relig_income
```

```{r}
relig_income %>% 
  pivot_longer(cols = !religion, names_to = "income", values_to = "count")
```

- pivot_wider(data, names_from, values_from)

```{r}
fish_encounters
```

```{r}
fish_encounters %>% 
  pivot_wider(names_from = station, values_from = seen)
```

- separate(data, col, into, sep) 

```{r}
df <- data.frame(x = c(NA, "x.y", "x.z", "y.z"))
df
df %>% separate(x, c("A", "B"))

# use regular expression in separator.
df %>% separate(x, c("A", "B"), sep = '\\.')
```

- unite()

```{r}
df_sep <- df %>% separate(x, c("A", "B"))

# Different separators.
df_sep %>% unite(x, A, B)

# don't need to use regular express for the separator
df_sep %>% unite(x, A, B, sep = ".")
```


# $13.4 Mutating joins

Compare the joins in dplyr and SQL.


dplyr	                                      SQL
-------------------------   --------------------------------------
inner_join(x, y, by = "z")	SELECT * FROM x INNER JOIN y USING (z)
left_join(x, y, by = "z")	  SELECT * FROM x LEFT OUTER JOIN y USING (z)
right_join(x, y, by = "z")	SELECT * FROM x RIGHT OUTER JOIN y USING (z)
full_join(x, y, by = "z")	  SELECT * FROM x FULL OUTER JOIN y USING (z)



# $13.7 Set operations

- intersect(x, y): return only obs. in both x and y.

- union(x, y): return unique obs. in x and y.

- setdiff(x, y): return obs. in x, but not in y.


# Exercise:

Compare the sheet **PBI New** and **SAS New** of Excel file: **Test PBI and SAS new and endorsement 20220214.xlsx**

- Policy number is the primary key for each sheet. Find out the difference records between the two sheets.

- use join function and setdiff function to see if you can get the same results.

 














