---
title: 'Lesson 4: Review Chapter 5 and Chapter 8'
author: "Hao Wang"
date: "2/2/2022"
output:
  pdf_document: default
---

# 0. Load libraries

```{r setup, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(rlang)

# Load NYC flight dataset 
library(nycflights13)
```

# 1. Review Chapter 5 Exercises

## \S 5.2.4

1.  Find all flights that

1.1 Had an arrival delay of two or more hours

```{r}
flights %>% filter(arr_delay >= 120)
```

2.2 The flights that flew to Houston are those flights where the destination (dest) is either "IAH" or "HOU".

```{r}
flights %>% filter(dest %in% c("IAH", "HOU"))
```

## \S 5.5.2

Q 2. Compare air_time with arr_time - dep_time. What do you expect to see? What do you see? What do you need to do to fix it?

**Expect**: air_time = arr_time - dep_time

Check it

```{r}
flights_airtime <-
  mutate(flights,
    dep_time_mins = (dep_time %/% 100 * 60 + dep_time %% 100) %% 1440,
    arr_time_mins = (arr_time %/% 100 * 60 + arr_time %% 100) %% 1440,
    air_time_diff = air_time - arr_time_mins + dep_time_mins
  )
```

```{r}
ggplot(flights_airtime, aes(x = air_time_diff)) +
  geom_histogram(binwidth = 1)
```

**Explanation:** The flights data does not contain the variables TaxiIn, TaxiOff, WheelsIn, and WheelsOff. It appears that the air_time variable refers to flight time, which is defined as the time between wheels-off (take-off) and wheels-in (landing). But the flight time does not include time spent on the runway taxiing to and from gates. With this new understanding of the data, the relationship between air_time, arr_time, and dep_time is air_time \<= arr_time - dep_time, supposing that the time zones of arr_time and dep_time are in the same time zone.

## \S 5.7.1

Q 6. Look at each destination. Can you find flights that are suspiciously fast? (i.e. flights that represent a potential data entry error). Compute the air time of a flight relative to the shortest flight to that destination. Which flights were most delayed in the air?

**Answers**: standardizing variables with the mean and variance, we could use the median as a measure of central tendency and the interquartile range (IQR) as a measure of spread. The median and IQR are more resistant to outliers than the mean and standard deviation.

$$standardized(x) = \frac{x - median(x)}{IQR(x)}$$

```{r}
standardized_flights <- flights %>%
  filter(!is.na(air_time)) %>%
  group_by(dest, origin) %>%
  mutate(
    air_time_median = median(air_time),
    air_time_iqr = IQR(air_time),
    n = n(),
    air_time_standard = (air_time - air_time_median) / air_time_iqr)
```

```{r}
ggplot(standardized_flights, aes(x = air_time_standard)) +
  geom_density()
```

```{r}
standardized_flights %>%
  arrange(air_time_standard) %>%
  select(
    carrier, flight, origin, dest, month, day, air_time,
    air_time_median, air_time_standard
  ) %>%
  head(10) %>%
  print(width = Inf)
```

Check the ground speed of flights.

```{r}
flights %>%
  mutate(mph = distance / (air_time / 60)) %>%
  ggplot(aes(x = mph)) +
  geom_histogram(binwidth = 10)
```

The fastest flight is

```{r}
flights %>%
  mutate(mph = distance / (air_time / 60)) %>%
  arrange(desc(mph)) %>%
  select(mph, flight, carrier, flight, month, day, dep_time) %>%
  head(5)
```

The most delay flight compare to the fastest flight in same des and arr.

air time comparing to the fastest flight on the route.

```{r}
air_time_delayed <-
  flights %>%
  group_by(origin, dest) %>%
  mutate(
    air_time_min = min(air_time, na.rm = TRUE),
    air_time_delay = air_time - air_time_min,
    air_time_delay_pct = air_time_delay / air_time_min * 100
  )
```

```{r}
air_time_delayed %>%
  arrange(desc(air_time_delay)) %>%
  select(
    air_time_delay, carrier, flight,
    origin, dest, year, month, day, dep_time,
    air_time, air_time_min
  ) %>%
  head() %>%
  print(width = Inf)
```

Q 8. For each plane, count the number of flights before the first delay of greater than 1 hour.

If we use the dep_delay, here is the code.

```{r}
flights %>%
  # sort in increasing order
  select(tailnum, year, month, day, dep_delay) %>%
  filter(!is.na(dep_delay)) %>%
  arrange(tailnum, year, month, day) %>%
  group_by(tailnum) %>%
  # cumulative number of flights delayed over one hour
  mutate(cumulative_hr_delays = cumsum(dep_delay > 60)) %>% #head(20)
  # count the number of flights == 0
  summarise(total_flights = sum(cumulative_hr_delays < 1)) %>%
  arrange(desc(total_flights))
```

# 2. Chapter 8. Workflow: projects in R Studio

-   where is your working directory?

```{r, echo=TRUE, results='hide'}
getwd()
```

-   Use RStudio Project to control the working folder and other folders

-   in the project folder, you can have sub-folders: code, raw_data, output, etc.

-   use relative path "./" (current path) and "../" (the parent path)

```{r}
ggplot(diamonds, aes(carat, price)) + 
  geom_hex()

ggsave("./output/diamonds.pdf")

write_csv(diamonds, "./output/diamonds.csv")
```
