---
title: "Study Plan for 'R for Data Science' "
author: "Hao Wang"
date: "1/11/2022"
output: pdf_document
---

**Notes**

This is a study plan for the guided self-study of [R4DS](http://r4ds.had.co.nz/).

**References**

Hadley Wickham and Garrett Grolemund: *R for Data Science: Import, Tidy, Transform, Visualize, and Model Data*. 1st ed., O'Reilly Media, 2017.

## Course Time line

Here is the study plan for each week. We will not study the Model Section together, but you are welcome to study it by yourself.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(readxl)
```

```{r weekly plan, echo=FALSE}
weekly_plan <- readxl::read_xlsx(path = "weekly time plan.xlsx") %>% 
  select(-`Preparation need from instructor`)

```

```{r, echo=FALSE}
knitr::kable(weekly_plan, "pipe")
```

We will have a 45 \~ 60 minutes tutorial session every week. During the tutorial session, I will

-   Go through some content in the book.

-   Discussion assignment/code

The expected self-study time for students will be 2 \~ 3 hours each week. And, if it possible, please send in your homework a day before the tutorial session.
